---
title: "Math598FinalProject"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data = read.csv('/Users/dr.z/Desktop/Math 598 Final Project/chart.csv')
#View(data)
```

```{r}
data$DateTime = as.POSIXct(data$DateTime)
plot(data$DateTime, data$Users, type='l', col='blue')
```

Work with data from 2015 to 2020 so we can make a more accurate model (omitting the many gaps in the early data and the erratic spikes after COVID).

```{r}
data2 = data[data$DateTime >= "2015-01-01" & data$DateTime < "2020-03-10",]
plot(data2$DateTime, data2$Users, type='l', col='blue')
```

Linear trend in the data, so we will take a log transformation.

```{r}
logUsers2 = log(data2$Users)
plot(data2$DateTime, logUsers2, type='l', col='blue')
```

Plot ACF and PACF of log transformed data.

```{r}
acf(logUsers2, na.action = na.pass)
pacf(logUsers2, na.action = na.pass)
```

The ACF decays slowly, so we will take the first difference.

```{r}
diff_Users2 = diff(logUsers2)
acf(diff_Users2, na.action = na.pass)
pacf(diff_Users2, na.action = na.pass)
```

After taking the first difference, we can see that there is non stationarity in the seasonal component. We will take a seasonal difference to account for this with s = 7.

```{r}
diff2_Users2 = diff(diff_Users2, 7)
acf(diff2_Users2, na.action = na.pass)
pacf(diff2_Users2, na.action = na.pass)
```

From the ACF and PACF, we will first determine the seasonal component of the SARIMA model. The ACF cuts off after lag P = 1 and the PACF tails off, which suggests that the seasonal component is AR(P) with P = 1. Hence, the purely seasonal model is SARIMA(1,1,0)$_{7}$.

We will now determine the non seasonal component of the model. Similarly to the seasonal component, we see that the ACF cuts off after lag p = 1 and the PACF tails off, which suggests that the non seasonal component is AR(p) with p = 1. Hence, the no seasonal model is SARIMA(1,1,0)$_{7}$.

We combine both components to obtain the multiplicative model SARIMA(1,1,0)x(1,1,0)$_{7}$. The diagnostics of the model are included below.

```{r}
library(astsa)
library(zoo)
as.ts(logUsers2)
logUsers_no_na = na.omit(logUsers2)
sarima(logUsers_no_na, 1, 1, 0, P = 1, D = 1, Q = 0, S = 7)
```

Forecasting ahead for five days (data is zoomed in). This is projected Steam users if COVID didn't occur.

```{r}
sarima.for(logUsers_no_na, 5, 1,1,0, 1,1,0,7)
```

```{r}
library(astsa)
library(zoo)
as.ts(data2$Users)
data2_no_na = na.omit(data2$Users)
sarima(data2_no_na, 1, 1, 0, P = 1, D = 1, Q = 0, S = 7)
```

```{r}
sarima.for(data2_no_na, 5, 1,1,0, 1,1,0,7)
```

\newpage

Building a model with the COVID data.

```{r}
data3 = data[data$DateTime >= "2015-01-01" & data$DateTime <= "2021-12-02",]
plot(data3$DateTime, data3$Users, type='l', col='blue')
```

```{r}
logUsers3 = log(data3$Users)
plot(data3$DateTime, logUsers3, type='l', col='blue')
```

```{r}
acf(logUsers3, na.action = na.pass)
pacf(logUsers3, na.action = na.pass)
```

```{r}
diff_Users3 = diff(logUsers3)
acf(diff_Users3, na.action = na.pass)
pacf(diff_Users3, na.action = na.pass)
```

```{r}
diff2_Users3 = diff(diff_Users3, 7)
acf(diff2_Users3, na.action = na.pass)
pacf(diff2_Users3, na.action = na.pass)
```

```{r}
library(astsa)
library(zoo)
as.ts(data3$Users)
Users3_no_na = na.omit(data3$Users)
sarima(Users3_no_na, 1, 1, 0, P = 1, D = 1, Q = 0, S = 7)
```

```{r}
sarima.for(Users3_no_na, 5, 1,1,0, 1,1,0,7)
```

```{r}
plot(Users3_no_na, type='l'); spectrum(Users3_no_na, log='no')
```


