---
title: "Math598FinalProject"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data = read.csv('/Users/dr.z/Desktop/Math 598 Final Project/chart.csv')
#View(data)
```

```{r}
data$DateTime = as.POSIXct(data$DateTime)
plot(data$DateTime, data$Users, type='l', col='blue')
```

The data has a clear trend, so we will take a log transformation.

```{r}
logUsers = log(data$Users)
plot(data$DateTime, logUsers, type='l', col='blue')
```

With the trend eliminatedd, we will now look at the ACF and PACF of the log transformed data.

```{r}
acf(logUsers, na.action = na.pass)
pacf(logUsers, na.action = na.pass)
```

The ACF decays slowly, so we will take the first difference. We can see that there is a seasonal component (?) every 7 lags in the ACF.

```{r}
diff_Users = diff(logUsers)
acf(diff_Users, na.action = na.pass)
pacf(diff_Users, na.action = na.pass)
```

There is nonstationarity in the seasonal component, so we will take a seasonal difference with s = 12 (? might change later).

```{r}
diff2_Users = diff(diff_Users, 12)
acf(diff2_Users, na.action = na.pass)
pacf(diff2_Users, na.action = na.pass)
```

```{r}
diff3_Users = diff(diff2_Users, 12)
acf(diff3_Users, na.action = na.pass)
pacf(diff3_Users, na.action = na.pass)
```

```{r}
sarima(x, 1, 1, 0, P = 1, D = 1, Q = 0, S = 12)
```


